---
title: "change_of_symptom_over_time"
author: "Jianyu"
date: '2022-06-02'
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
library(grid)
library(gridExtra)
set.seed(42)
```

Difference from the aerosol_mts_plot.Rmd:
limit to those with some detection in aerosol samples and lump the two Omicrons together
```{r include=FALSE}
# Theme for display parameters

scale_color_variant <- scale_color_manual(values = c("A/O"= "#984ea3",
                                                     "Alpha"="#386cb0",
                                                     "Delta"="#7fc97f",
                                                     "Omicron" = "#ef3b2c"))
scale_shape_variant <- scale_shape_manual(values = c("A/O" = 15,
                                                     "Alpha"=16,
                                                     "Delta"=17,
                                                     "Omicron" = 18))

scale_color_variant3 <- scale_color_manual(values = c("A/O" = "#984ea3",
                                                     "Alpha"="#386cb0",
                                                     "Delta"="#7fc97f",
                                                     "BA.1" = "#a6cee3",
                                                     "BA.2" = "#662506"))
scale_shape_variant3 <- scale_shape_manual(values = c("A/O" = 15,
                                                     "Alpha"=16,
                                                     "Delta"=17,
                                                     "BA.1" = 18,
                                                     "BA.2" = 19))

```

```{r}
spr_before_raw <- read.csv(file = "../working_files/manuscript_1_all_61_nonpaired_pcr_sx_spr.csv")

spr_raw <- read.csv(file = "../working_files/nonpaired_pcr_sx_spr.csv")

spr <- spr_raw %>% mutate(variant2 = if_else(seq_variant %in% c("Omicron BA.1.1","Omicron BA.1", "Omicron BA.2"), "Omicron", seq_variant))
spr <- spr %>% mutate(variant3 = ifelse(seq_variant %in% c("Omicron BA.1.1","Omicron BA.1"), "BA.1", ifelse(seq_variant=="Omicron BA.2", "BA.2", seq_variant)))

spr_before <- spr_before_raw %>%
  mutate(subject_id = as.factor(study_id),
         days_past_onset = as.factor(days_past_onset),
         variant2 = if_else(S.gene.dropout == "TRUE", "Alpha", "A/O")) %>% 
  mutate(seq_variant = variant2,
         variant3 = variant2) 
length(unique(spr_before$subject_id))


before_sub <- spr_before %>% 
  select(subject_id, sample_date, days_past_onset, variant2, variant3, seq_variant, nose_run,nose_stuf,sneeze,throat_sr,earache,malaise,headache,mj_ache,sw_fever_chill,nausea,loa,vomit,diarrhea,chest_tight,sob,cough,ur,lr,gi,sys) %>%
  unique()

spr <- spr %>%
  mutate(subject_id = as.factor(subject_id),
         days_past_onset = as.factor(days_past_onset))

now_sub <- spr %>% 
  select(subject_id, sample_date, days_past_onset, variant2, variant3, seq_variant, nose_run,nose_stuf,sneeze,throat_sr,earache,malaise,headache,mj_ache,sw_fever_chill,nausea,loa,vomit,diarrhea,chest_tight,sob,cough,ur,lr,gi,sys) %>% 
  unique()

symptom_total <- bind_rows(now_sub,before_sub)

symptom_total <- symptom_total %>% 
  mutate(variant2 = factor(variant2, 
                          levels = c("A/O", "Alpha", "Delta", "Omicron"),
                          labels = c("A/O", "Alpha", "Delta", "Omicron"))) %>% 
  mutate(variant3 = factor(variant3, 
                          levels = c("A/O", "Alpha", "Delta", "BA.1","BA.2"),
                          labels = c("A/O", "Alpha", "Delta",  "BA.1","BA.2")))

symptom_total <- symptom_total %>% mutate(omicron = if_else(variant2 == "Omicron", "Yes", "No"))
```
individual symptom by variants over time with significant comparison
```{r}
stat_box_data <- function(y, upper_limit = 9.085) {
  data.frame(
    y = -0.02 * upper_limit,
    label = paste0('n=', length(y), '\n')
  )
}

symptom_total <- symptom_total %>% mutate(sample_date2 = as.Date(sample_date, origin="1970-01-01"))
variables <- colnames(symptom_total[,7:26])
variables_names <- c("Running nose", "Stuffy Nose", "Sneeze", "Sore throat", "Earache", "Malaise", "Headache", "Muscle ache", "Sweat/fever/chill","Nausea", "Loss of appetite", "Vomit", "Diarrhea", "Chest tightness", "Short of breath", "Cough", "Upper-respiratory", "Lower-respiratory", "Gastrointestinal", "Systemic")


compare <- list(c("A/O","Alpha"),c("Alpha","Delta"), c("Delta","BA.1"),c("BA.1","BA.2"),
               c("A/O","Delta"),c("Delta","BA.2"), 
               c("Alpha","BA.1"),
               c("A/O","BA.1"),
               c("Alpha","BA.2"),
               c("A/O","BA.2"))

label_compare = c(3,3,3,3,
                  3.5,3.5,
                  4,
                  4.5,
                  5,
                  5.5)

#The comparisons need to be defined individually and change everytime we run the code for new dataset
#This will be defined based on the results from line 198-219
compare_1 <- list(c("A/O","BA.1"), c("A/O","BA.2"))
label_compare_1 = c(3.5,4.5)

compare_2 <- list(c("Delta","BA.2"), c("A/O","BA.2"))
label_compare_2 = c(3.5,4.5)

compare_3 <- list(c("A/O","BA.1"), c("A/O","BA.2"))
label_compare_3 = c(3.5,4.5)

compare_4 <- list(c("A/O","BA.1"), c("Delta","BA.1"))
label_compare_4 = c(3.5,4.5)

compare_5 <- list(c("A/O","Alpha"))
label_compare_5 = c(3.5)

compare_6 <- list(c("A/O","Alpha"), c("A/O","BA.2"))
label_compare_6 = c(3.5,4.5)

compare_7 <- list(c("Delta","Alpha"), c("Delta","BA.1"))
label_compare_7 = c(3.5,4.5)

compare_8 <- list(c("A/O","Alpha"), c("A/O","BA.1"))
label_compare_8 = c(3.5,4.5)

compare_9 <- list(c("A/O","Alpha"), c("A/O","BA.1"),c("A/O","BA.2"))
label_compare_9 = c(3.5,4.5,5.5)

compare_10 <- list(c("A/O","Alpha"), c("A/O","BA.2"))
label_compare_10 = c(3.5,4.5)

compare_11 <- list(c("A/O","Alpha"),c("Alpha","Delta"), c("Delta","BA.1"),
                   c("Delta","BA.2"),c("A/O","BA.1"),
                   c("A/O","BA.2"))
label_compare_11 = c(3.5,3.5,3.5,
                     4.5,
                     5.5,
                     6.5)
compare_12 <- list()
label_compare_12 = c()

compare_13 <- list(c("A/O","Alpha"), c("A/O","BA.1"))
label_compare_13 = c(3.5,4.5)

compare_14 <- list()
label_compare_14 = c()

compare_15 <- list(c("A/O","BA.1"))
label_compare_15 = c(3.5)

compare_16 <- list(c("A/O","BA.1"), c("A/O","BA.2"))
label_compare_16 = c(3.5,4.5)

symptom_by_variant_by_day3 <- function(virus, symptom, symptom_name, comparison, label_comparison){
  plot <- symptom_total %>% 
  ggplot(aes(x = eval(parse(text=virus)), y = eval(parse(text=symptom))))+
  geom_boxplot(aes(color=eval(parse(text=virus))),show.legend = FALSE)+
  stat_summary(
    fun.data = stat_box_data, 
    geom = "text", 
    hjust = 0.5,
    vjust = 0.9,
    size=3,
    position = position_dodge(0.9)
  ) + 
  stat_compare_means(comparisons = comparison,size=3, label ="p.signif", label.y = label_comparison)+
  stat_compare_means(label.y=7.5, label.x = "Alpha", size=3)+
  coord_cartesian(ylim = c(-1, 8), expand = TRUE)+
  scale_y_continuous(breaks=c( 0,1,2,3))+
  theme_bw()+
  scale_shape_variant3+
  scale_color_variant3+
    labs(x= " ", 
       y=symptom_name,
       color = "Variant",
       shape = "Variant")+
  theme(axis.text.x=element_text(size = 10, face='bold', angle = 0),
        axis.text.y=element_text(size = 12, face='bold'),
        axis.title.y=element_text(size=12, face='bold'),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16, face='bold'))
  return(plot)
}
```

```{r eval=FALSE, include=FALSE}
#display all the pairwise comparison to find out which to display in the final plot
all_individual_symptom<- ggarrange(
  symptom_by_variant_by_day3("variant3", variables[1], variables_names[1], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[2], variables_names[2], compare, label_compare),
  symptom_by_variant_by_day3("variant3", variables[3], variables_names[3], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[4], variables_names[4], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[5], variables_names[5], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[6], variables_names[6], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[7], variables_names[7], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[8], variables_names[8], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[9], variables_names[9], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[10], variables_names[10], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[11], variables_names[11], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[12], variables_names[12], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[13], variables_names[13], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[14], variables_names[14], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[15], variables_names[15], compare, label_compare), 
  symptom_by_variant_by_day3("variant3", variables[16], variables_names[16], compare, label_compare), 
                      ncol=4, nrow=4)
all_individual_symptom
```

```{r}
individual_symptom<- ggarrange(
  symptom_by_variant_by_day3("variant3", variables[1], variables_names[1], compare_1, label_compare_1), 
  symptom_by_variant_by_day3("variant3", variables[2], variables_names[2], compare_2, label_compare_2),
  symptom_by_variant_by_day3("variant3", variables[3], variables_names[3], compare_3, label_compare_3), 
  symptom_by_variant_by_day3("variant3", variables[4], variables_names[4], compare_4, label_compare_4), 
  symptom_by_variant_by_day3("variant3", variables[5], variables_names[5], compare_5, label_compare_5), 
  symptom_by_variant_by_day3("variant3", variables[6], variables_names[6], compare_6, label_compare_6), 
  symptom_by_variant_by_day3("variant3", variables[7], variables_names[7], compare_7, label_compare_7), 
  symptom_by_variant_by_day3("variant3", variables[8], variables_names[8], compare_8, label_compare_8), 
  symptom_by_variant_by_day3("variant3", variables[9], variables_names[9], compare_9, label_compare_9), 
  symptom_by_variant_by_day3("variant3", variables[10], variables_names[10], compare_10, label_compare_10), 
  symptom_by_variant_by_day3("variant3", variables[11], variables_names[11], compare_11, label_compare_11), 
  symptom_by_variant_by_day3("variant3", variables[12], variables_names[12], compare_12, label_compare_12), 
  symptom_by_variant_by_day3("variant3", variables[13], variables_names[13], compare_13, label_compare_13), 
  symptom_by_variant_by_day3("variant3", variables[14], variables_names[14], compare_14, label_compare_14), 
  symptom_by_variant_by_day3("variant3", variables[15], variables_names[15], compare_15, label_compare_15), 
  symptom_by_variant_by_day3("variant3", variables[16], variables_names[16], compare_16, label_compare_16), 
                      ncol=4, nrow=4)

ggsave(individual_symptom,filename="../output/si_fig9a_box_sx_all.png",device="png",width=11,height=7.77,units="in")

```

#symptom scores by variants over time with significant comparison
```{r}
compare_17 <- list(c("A/O","BA.1"), c("A/O","BA.2"))
label_compare_17 = c(15.5, 17)

compare_18 <- list(c("A/O","BA.1"), c("A/O","BA.2"))
label_compare_18 = c(15, 16.5)

compare_19 <- list(c("A/O","Alpha"),c("Alpha","Delta"), c("Delta","BA.1"),
                   c("Delta","BA.2"),c("A/O","BA.1"),
                   c("A/O","BA.2"))
label_compare_19 = c(14.5,14.5,14.5,
                     15.7,
                     17,
                     18.2)

compare_20 <- list(c("A/O","Alpha"),c("A/O","BA.1"), c("Delta","BA.1"))
label_compare_20 = c(15,16.5)

symptom_by_variant_by_day4 <- function(virus, symptom, symptom_name, comparison, label_comparison, note_text){
  plot <- symptom_total %>% 
  ggplot(aes(x = eval(parse(text=virus)), y = eval(parse(text=symptom))))+
  geom_boxplot(aes(color=eval(parse(text=virus))),show.legend = TRUE)+
  scale_y_continuous(breaks=c( 0,5,10,15))+
  coord_cartesian(ylim = c(-1, 20), expand = TRUE)+
  stat_summary(
    fun.data = stat_box_data, 
    geom = "text", 
    hjust = 0.5,
    vjust = 0.9,
    size=3,
    position = position_dodge(0.9)
  ) +
  stat_compare_means(label.y=19.8, label.x = "Alpha", size=3)+
  stat_compare_means(comparisons = comparison,size=3, label ="p.signif", label.y = label_comparison)+
  ggtitle(note_text)+
  theme_bw()+
  scale_shape_variant3+
  scale_color_variant3+
    labs(x= " ", 
       y=symptom_name,
       color = "Variant",
       shape = "Variant")+
  theme(axis.text.x=element_text(size = 10, face='bold', angle = 0),
        axis.text.y=element_text(size = 12, face='bold'),
        axis.title.y=element_text(size=12, face='bold'),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16, face='bold'),
        plot.title = element_text(size = 20, face = "bold"))+
   guides(shape = guide_legend(override.aes = list(size = 5)))+
  theme(legend.position="bottom")
  return(plot)
}

```

```{r eval=FALSE, include=FALSE}
#display all the pairwise comparison to find out which to display in the final plot
all_figure4 <- ggarrange(
  symptom_by_variant_by_day4("variant3", variables[17], variables_names[17], compare, label_compare,"a"), 
  symptom_by_variant_by_day4("variant3", variables[18], variables_names[18], compare, label_compare,"b"), 
  symptom_by_variant_by_day4("variant3", variables[19], variables_names[19], compare, label_compare,"c"), 
  symptom_by_variant_by_day4("variant3", variables[20], variables_names[20], compare, label_compare,"d"), 
  nrow=2, ncol=2,
  legend = "none")
all_figure4
```

```{r eval=FALSE, include=FALSE}
figure4 <- ggarrange(
  symptom_by_variant_by_day4("variant3", variables[17], variables_names[17], compare_17, label_compare_17,"a"), 
  symptom_by_variant_by_day4("variant3", variables[18], variables_names[18], compare_18, label_compare_18,"b"), 
  symptom_by_variant_by_day4("variant3", variables[19], variables_names[19], compare_19, label_compare_19,"c"), 
  symptom_by_variant_by_day4("variant3", variables[20], variables_names[20], compare_20, label_compare_20,"d"), 
  nrow=2, ncol=2,
  legend = "none")
figure4
```

##for cough data
```{r}
spr <- spr_raw %>% mutate(variant2 = if_else(seq_variant %in% c("Omicron BA.1.1","Omicron BA.1", "Omicron BA.2"), "Omicron", seq_variant))
spr <- spr %>% mutate(variant3 = ifelse(seq_variant %in% c("Omicron BA.1.1","Omicron BA.1"), "BA.1", ifelse(seq_variant=="Omicron BA.2", "BA.2", seq_variant)))

spr_before <- spr_before_raw %>%
  filter(sample_type %in% c("G-II Fine Aerosol")) %>% 
  mutate(subject_id = as.factor(study_id),
         days_past_onset = as.factor(days_past_onset),
         variant2 = if_else(S.gene.dropout == "TRUE", "Alpha", "A/O")) %>% 
  mutate(seq_variant = variant2,
         variant3 = variant2)
length(unique(spr_before$subject_id))

before_sub <- spr_before %>% 
  select(subject_id, sample_date, sample_type, days_past_onset, variant2, variant3, seq_variant, cough_count)

spr <- spr %>%
  filter(sample_type %in% c("G-II Fine Aerosol")) %>% 
  mutate(subject_id = as.factor(subject_id),
         days_past_onset = as.factor(days_past_onset))

now_sub <- spr %>% 
  select(subject_id, sample_date, sample_type, days_past_onset, variant2, variant3, seq_variant, cough_count) 

breath_total <- bind_rows(now_sub,before_sub)

breath_total <- breath_total %>% 
  mutate(variant2 = factor(variant2, 
                          levels = c("A/O", "Alpha", "Delta", "Omicron"),
                          labels = c("A/O", "Alpha", "Delta", "Omicron"))) %>% 
  mutate(variant3 = factor(variant3, 
                          levels = c("A/O", "Alpha", "Delta", "BA.1","BA.2"),
                          labels = c("A/O", "Alpha", "Delta",  "BA.1","BA.2")))

breath_total <- breath_total %>% mutate(omicron = if_else(variant2 == "Omicron", "Yes", "No"))
```

#create the cough count during sample collection by variant over time
```{r}
breath_total <- breath_total %>% mutate(sample_date2 = as.Date(sample_date, origin="1970-01-01"))

cough_by_variant_by_day1 <- breath_total %>% 
  ggplot(aes(x = sample_date2, y = cough_count))+
  geom_point(aes(color=variant2, shape = variant2))+
  scale_y_continuous(breaks=c( 0,10,20,30,40,50,60,70))+
  coord_cartesian(ylim = c(-5, 90), expand = TRUE)+
  theme_bw()+
  scale_shape_variant+
  scale_color_variant+
  scale_x_date(date_labels = "%b %Y", breaks = "4 month")+
  labs(x= " ", 
       y="Number of coughs",
       color = "Variant",
       shape = "Variant")+
  ggtitle("a")+
  theme(axis.text.x=element_text(size = 10, face='bold', angle = 0),
        axis.text.y=element_text(size = 10),
        axis.title.y=element_text(size=12, face='bold'),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16, face='bold'),
        plot.title=element_text(size=20, face="bold"))+
   guides(shape = guide_legend(override.aes = list(size = 5)))+
  theme(legend.position="bottom")
cough_by_variant_by_day1
```

#create the pairwise comparisons for figure 2
```{r}
compare_all <- list(c("A/O", "Alpha"), c("Delta", "Omicron"), c("Alpha","Delta"), c("Omicron", "Alpha"), c("Delta", "A/O"), c("Omicron", "A/O"))
label_compare_all <- c(40, 50, 58,67,74,80)

compare<- list(c("Alpha","Delta"), c("Omicron", "Alpha"), c("Delta", "A/O"), c("Omicron", "A/O"))
label_compare = c(58,67,74,80)
```

```{r eval=FALSE, include=FALSE}
#display all the pairwise comparison to find out which to display in the final plot
all_cough_by_variant_by_day2 <- breath_total%>% 
  ggplot(aes(x = variant2, y = cough_count))+
  geom_boxplot(aes(color = variant2),show.legend = FALSE)+
  labs(x= " ", 
       y=" ")+
  stat_summary(
    fun.data = stat_box_data, 
    geom = "text", 
    hjust = 0.5,
    vjust = 0.9,
    size=3,
    position = position_dodge(0.9)
  ) + 
  scale_y_continuous(breaks=c( 0,10,20,30,40,50,60,70))+
  coord_cartesian(ylim = c(-5, 90), expand = TRUE)+
  stat_compare_means(comparisons = compare_all,size=3, label ="p.signif", label.y = label_compare_all)+
  stat_compare_means(label.y=89, label.x = "A/O", size=3)+
  theme_bw()+
  scale_shape_variant+
  scale_color_variant+
  ggtitle("b")+
  theme(axis.text.x=element_text(size = 10, face='bold', angle = 00),
        axis.text.y=element_text(size = 10),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16, face='bold'),
        plot.title=element_text(size=20, face="bold"))

all_cough_by_variant_by_day2
```

```{r}
cough_by_variant_by_day2 <- breath_total%>% 
  ggplot(aes(x = variant2, y = cough_count))+
  geom_boxplot(aes(color = variant2),show.legend = FALSE)+
  labs(x= " ", 
       y=" ")+
  stat_summary(
    fun.data = stat_box_data, 
    geom = "text", 
    hjust = 0.5,
    vjust = 0.9,
    size=3,
    position = position_dodge(0.9)
  ) + 
  scale_y_continuous(breaks=c( 0,10,20,30,40,50,60,70))+
  coord_cartesian(ylim = c(-5, 90), expand = TRUE)+
  stat_compare_means(comparisons = compare,size=3, label ="p.signif", label.y = label_compare)+
  stat_compare_means(label.y=89, label.x = "A/O", size=3)+
  theme_bw()+
  scale_shape_variant+
  scale_color_variant+
  ggtitle("b")+
  theme(axis.text.x=element_text(size = 10, face='bold', angle = 00),
        axis.text.y=element_text(size = 10),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16, face='bold'),
        plot.title=element_text(size=20, face="bold"))

cough_by_variant_by_day2
```

combine the plots
```{r}

cough<- ggarrange(cough_by_variant_by_day1, cough_by_variant_by_day2,
                  ncol = 2, nrow=1,
                  widths = c(2,1),
                  common.legend = T,
                  legend = "bottom",
                  align = "h")
cough

symptom <- ggarrange(symptom_by_variant_by_day4("variant3", variables[17], variables_names[17], compare_17, label_compare_17,"c"),
                      symptom_by_variant_by_day4("variant3", variables[18], variables_names[18], compare_18, label_compare_18,"d"),
                      symptom_by_variant_by_day4("variant3", variables[19], variables_names[19], compare_19, label_compare_19,"e"),
                      symptom_by_variant_by_day4("variant3", variables[20], variables_names[20], compare_20, label_compare_20,"f"),

                      font.label = list(size = 20, face = "bold"),
                      ncol = 2, nrow=2,
                      common.legend = T,
                      legend = "none",
                      align = "h")
```


```{r}
blank <- grid.rect(gp=gpar(col="white"))
total <- grid.arrange(cough, blank, symptom,
                      ncol=1, nrow=3,
                      heights = c(6,1,9))
total
ggsave(total,filename="../output/si_fig8a_box_comp_sx.png",device="png",dpi=300, width=10,height=11,units="in")
# ggsave(total,filename="../output/tiff/extend_figure_8a_cough_symptom_score_change.tiff",device="tiff",dpi=600, width=10,height=11,units="in",compression="lzw")
# used to be in main manuscript, but got moved to SI
```
