---
title: "plots_by_variants_by_time"
author: "Jianyu"
date: "3/9/2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
library(grid)
library(gridExtra)
set.seed(42)
```

Difference from the aerosol_mts_plot.Rmd:
limit to those with some detection in aerosol samples and lump the two Omicrons together
```{r include=FALSE}
# Theme for display parameters

scale_color_variant <- scale_color_manual(values = c("Ancestral/other" = "#984ea3",
                                                     "Alpha"="#386cb0",
                                                     "Delta"="#7fc97f",
                                                     "Omicron" = "#ef3b2c"))
scale_shape_variant <- scale_shape_manual(values = c("Ancestral/other" = 15,
                                                     "Alpha"=16,
                                                     "Delta"=17,
                                                     "Omicron" = 18))

scale_color_variant3 <- scale_color_manual(values = c("Ancestral/other" = "#984ea3",
                                                     "Alpha"="#386cb0",
                                                     "Delta"="#7fc97f",
                                                     "Omicron BA.1" = "#a6cee3",
                                                     "Omicron BA.2" = "#662506"))
scale_shape_variant3 <- scale_shape_manual(values = c("Ancestral/other" = 15,
                                                     "Alpha"=16,
                                                     "Delta"=17,
                                                     "Omicron BA.1" = 18,
                                                     "Omicron BA.2" = 19))

```

```{r}
spr_before <- read.csv(file = "../working_files/manuscript_1_all_61_nonpaired_pcr_sx_spr.csv")
spr_before <- spr_before %>% filter(days_past_onset<6|is.na(days_past_onset))

spr <- read.csv(file = "../working_files/nonpaired_pcr_sx_spr.csv")
spr <- spr %>% filter(days_past_onset<6|is.na(days_past_onset))

spr <- spr %>% mutate(variant2 = if_else(seq_variant %in% c("Omicron BA.1.1","Omicron BA.1", "Omicron BA.2"), "Omicron", seq_variant))
spr <- spr %>% mutate(variant3 = if_else(seq_variant %in% c("Omicron BA.1.1","Omicron BA.1"), "Omicron BA.1", seq_variant))
samples_of_interest <- c("G-II Fine Aerosol", "G-II Coarse Aerosol", "Midturbinate Swab")

spr_before <- spr_before %>% 
  mutate(log_quant = if_else(av_quant == 0, 0, log10(av_quant)),
         subject_id = as.factor(study_id),
         days_past_onset = as.factor(days_past_onset),
         period = "Before Sep 2021",
         variant2 = if_else(S.gene.dropout == "TRUE", "Alpha", "Ancestral/other")) %>% 
  mutate(seq_variant = variant2,
         variant3 = variant2)
length(unique(spr_before$subject_id))

before_sub <- spr_before %>% 
  filter(sample_type %in% samples_of_interest) %>% 
  select(subject_id, sample_date, av_quant, log_quant, sample_type, period, days_past_onset, variant2, variant3, seq_variant)

spr <- spr %>% 
  mutate(log_quant = if_else(av_quant == 0, 0, log10(av_quant)),
         subject_id = as.factor(subject_id),
         days_past_onset = as.factor(days_past_onset),
         period = "After Sep 2021")

now_sub <- spr %>% filter(sample_type %in% samples_of_interest) %>% 
  select(subject_id, sample_date, av_quant, log_quant, sample_type, period, days_past_onset, variant2, variant3, seq_variant)

breath_mts_total <- bind_rows(now_sub,before_sub)

breath_mts_total <- breath_mts_total %>% 
  mutate(variant2 = factor(variant2, 
                          levels = c("Ancestral/other", "Alpha", "Delta", "Omicron"),
                          labels = c("Ancestral/other", "Alpha", "Delta", "Omicron"))) %>% 
  mutate(variant3 = factor(variant3, 
                          levels = c("Ancestral/other", "Alpha", "Delta", "Omicron BA.1","Omicron BA.2"),
                          labels = c("Ancestral/other", "Alpha", "Delta",  "Omicron BA.1","Omicron BA.2")))

breath_mts_total <- breath_mts_total %>% mutate(omicron = if_else(variant2 == "Omicron", "Yes", "No"))
```
prepare complete dataset and dataset limited data to at least one of the breath samples is not 0 per subject-sample_date:
```{r}
stat_box_data <- function(y, upper_limit = 9.085) {
    data.frame(
      y = -0.02 * upper_limit,
      label = paste0('n=', length(y), '\n')
    )
}

fine_complete <- breath_mts_total %>% 
  filter(sample_type == "G-II Fine Aerosol") %>%
  mutate(fine_av_quant = av_quant, fine_log_quant = log_quant) %>% 
  select(subject_id, fine_av_quant, fine_log_quant, days_past_onset, sample_date, period, variant2, omicron, variant3,seq_variant) 
fine_complete_sub <- fine_complete %>%filter(fine_av_quant !=0)  

coarse_complete <- breath_mts_total %>% 
  filter(sample_type == "G-II Coarse Aerosol") %>%
  mutate(coarse_av_quant = av_quant, coarse_log_quant = log_quant) %>% 
  select(subject_id, coarse_av_quant, coarse_log_quant, days_past_onset, sample_date, period, variant2, omicron, variant3, seq_variant)
coarse <- coarse_complete %>% select(subject_id, coarse_av_quant, coarse_log_quant, sample_date)
coarse_complete_sub <- coarse_complete %>%filter(coarse_av_quant !=0)

mts_complete <- breath_mts_total %>% 
  filter(sample_type == "Midturbinate Swab") %>%
  mutate(mts_av_quant = av_quant, mts_log_quant = log_quant) %>% 
  select(subject_id, mts_av_quant, mts_log_quant, days_past_onset, sample_date, period, variant2, omicron, variant3, seq_variant)

mts <- mts_complete %>% select(subject_id, mts_av_quant, mts_log_quant, sample_date) 

#create dataset limited to at least one of the breath samples is non-zero: breath_mts_total_wide1
breath_mts_total_wide1 <- fine_complete_sub %>% 
  full_join(coarse_complete_sub, by = c("subject_id", "sample_date")) %>% 
  mutate(coarse_av_quant = if_else(is.na(coarse_av_quant), 0, coarse_av_quant),
         coarse_log_quant = if_else(is.na(coarse_log_quant), 0, coarse_log_quant),
         fine_av_quant = if_else(is.na(fine_av_quant), 0, fine_av_quant),
         fine_log_quant = if_else(is.na(fine_log_quant), 0, fine_log_quant),
         days_past_onset = if_else(is.na(days_past_onset.y), days_past_onset.x, days_past_onset.y),
         period = if_else(is.na(period.y), period.x, period.y),
         variant2 = if_else(is.na(variant2.y), variant2.x, variant2.y),
         omicron = if_else(is.na(omicron.y), omicron.x, omicron.y),
         variant3 = if_else(is.na(variant3.y), variant3.x, variant3.y),
         seq_variant = if_else(is.na(seq_variant.y), seq_variant.x, seq_variant.y)) %>% 
  mutate(av_quant_total = coarse_av_quant + fine_av_quant)

breath_mts_total_wide1 <- breath_mts_total_wide1 %>% 
  mutate(av_quant_total_log10 = if_else((av_quant_total == 0), 0, log10(av_quant_total)))

breath_mts_total_wide1 <- breath_mts_total_wide1 %>% 
  left_join(mts, by = c("subject_id", "sample_date")) %>% 
  mutate(log_mts_coarse_ratio = log10(mts_av_quant/(coarse_av_quant+1)),
         log_mts_fine_ratio = log10(mts_av_quant/(fine_av_quant+1)),
         log_mts_total_ratio = log10(mts_av_quant/(av_quant_total+1)),
         mts_coarse_ratio =  mts_av_quant/(coarse_av_quant+1),
         mts_fine_ratio = mts_av_quant/(fine_av_quant+1),
         mts_total_ratio = mts_av_quant/(av_quant_total+1))

#create dataset including all the breath samples:breath_mts_total_wide2
breath_mts_total_wide2 <- fine_complete %>% 
  left_join(coarse, by = c("subject_id", "sample_date")) %>% 
  mutate(av_quant_total = coarse_av_quant + fine_av_quant)

breath_mts_total_wide2 <- breath_mts_total_wide2 %>% 
  mutate(av_quant_total_log10 = if_else((av_quant_total == 0), 0, log10(av_quant_total)))


breath_mts_total_wide2 <- breath_mts_total_wide2 %>% 
  left_join(mts, by = c("subject_id", "sample_date")) %>% 
  mutate(log_mts_coarse_ratio = log10(mts_av_quant/(coarse_av_quant+1)),
         log_mts_fine_ratio = log10(mts_av_quant/(fine_av_quant+1)),
         log_mts_total_ratio = log10(mts_av_quant/(av_quant_total+1)),
         mts_coarse_ratio =  mts_av_quant/(coarse_av_quant+1),
         mts_fine_ratio = mts_av_quant/(fine_av_quant+1),
         mts_total_ratio = mts_av_quant/(av_quant_total+1))

length(unique(breath_mts_total_wide2$subject_id))
```
create the EBA viral load by variant over time
```{r}
breath_mts_total_wide2 <- breath_mts_total_wide2 %>% mutate(sample_date2 = as.Date(sample_date, origin="1970-01-01"))

sample_plot_by_variant_by_day1 <- function(data, sample_value, name, legend_show, note_text){
plot <- data %>% 
  ggplot(aes(x = sample_date2, y = eval(parse(text=sample_value))))+
  geom_point(aes(color=variant2, shape = variant2), show.legend = legend_show)+
  coord_cartesian(ylim = c(0, 8), expand = TRUE)+
  theme_bw()+
  scale_shape_variant+
  scale_color_variant+
  scale_x_date(date_labels = "%b %Y", breaks = "4 month")+
  labs(x= " ", 
       y=paste(name, "log10(RNA)"),
       color = "Variant",
       shape = "Variant")+
  ggtitle(note_text)+
  theme(axis.text.x=element_text(size = 12, face='bold', angle = 0),
        axis.text.y=element_text(size = 12, face='bold'),
        axis.title.x=element_text(size=12, face='bold'),
        axis.title.y=element_text(size=12, face='bold'),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16, face='bold'),
        plot.title = element_text(size = 20, face = "bold"))+
   guides(shape = guide_legend(override.aes = list(size = 5)))+
  theme(legend.position="bottom")
return(plot) 
}
```
create the pairwise comparisons for figure 2
```{r}
compare_all <- list(c("Ancestral/other", "Alpha"), c("Delta", "Omicron"), c("Alpha","Delta"), c("Omicron", "Alpha"), c("Delta", "Ancestral/other"), c("Omicron", "Ancestral/other"))
label_compare_all <- c(8,9,10,11,12,13)

compare_for_fine_and_total <- list( c("Alpha", "Ancestral/other"), c("Delta", "Ancestral/other"), c("Omicron", "Ancestral/other"))
label_fine_and_total = c(8, 10, 12)
compare_for_coarse <- list( c("Alpha", "Ancestral/other"), c("Delta", "Ancestral/other"), c("Omicron", "Ancestral/other"), c("Alpha", "Omicron"))
label_coarse = c(8, 10, 14, 12)

sample_plot_by_variant <- function(data, sample_value, name, compare,label_y,note_text){
plot <- data %>% 
  ggplot(aes(x = variant2, y = eval(parse(text=sample_value))))+
  #geom_jitter(aes(color = variant2, alpha = 0.1))+
  geom_boxplot(aes(color = variant2),show.legend = FALSE)+
  labs(x= " ", 
       y=" ")+
  stat_summary(
    fun.data = stat_box_data, 
    geom = "text", 
    hjust = 0.5,
    vjust = 0.9,
    size=3,
    position = position_dodge(0.9)
  ) + 
  scale_y_continuous(breaks=c( 0, 2, 4, 6, 8))+
  coord_cartesian(ylim = c(-2, 18), expand = FALSE)+
  stat_compare_means(comparisons = compare,size=3, label ="p.signif", label.y = label_y)+
  stat_compare_means(label.y=16, label.x = "Alpha", size=3)+
  theme_bw()+
  scale_shape_variant+
  scale_color_variant+
  ggtitle(note_text)+
  theme(axis.text.x=element_text(size = 10, face='bold', angle = 00),
        axis.text.y=element_text(size = 12, face='bold'),
        axis.title.x=element_text(size=10, face='bold'),
        axis.title.y=element_text(size=12, face='bold'),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16, face='bold'),
        plot.title = element_text(size = 20, face = "bold"))
  
return(plot) 
}
```

```{r eval=FALSE, include=FALSE}
#display all the pairwise comparison to find out which to display in the final plot
all_full_sample_by_variant_plot <- ggarrange(sample_plot_by_variant(breath_mts_total_wide2,  "fine_log_quant", "Fine",compare_all,label_compare_all,"A"),
          sample_plot_by_variant(breath_mts_total_wide2, "coarse_log_quant", "Coarse",compare_for_coarse,label_compare_all,"B"),
          sample_plot_by_variant(breath_mts_total_wide2, "av_quant_total_log10", "Total",compare_all,label_compare_all,"C"),
          #labels = c("A", "B", "C"),
          #hjust=-4,vjust= 2,
          #ncol = 1, nrow=3,
          common.legend = T,
          legend = "bottom",
          align = "h")
all_full_sample_by_variant_plot
```

```{r}
full_sample_by_variant_plot <- ggarrange(sample_plot_by_variant(breath_mts_total_wide2,  "fine_log_quant", "Fine",compare_for_fine_and_total,label_fine_and_total,"A"),
          sample_plot_by_variant(breath_mts_total_wide2, "coarse_log_quant", "Coarse",compare_for_coarse,label_coarse,"B"),
          sample_plot_by_variant(breath_mts_total_wide2, "av_quant_total_log10", "Total",compare_for_fine_and_total,label_fine_and_total,"C"),
          #labels = c("A", "B", "C"),
          #hjust=-4,vjust= 2,
          #ncol = 1, nrow=3,
          common.legend = T,
          legend = "bottom",
          align = "h")

full_sample_by_variant_plot
#ggsave(full_sample_by_variant_plot,filename="../output/pairwise_test_for_figure1.png",device="png",width=11,height=7.77,units="in")
```

legend only
```{r}
## Function to extract legend
g_legend <- function(a.gplot){ 
    tmp <- ggplot_gtable(ggplot_build(a.gplot)) 
    leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box") 
    legend <- tmp$grobs[[leg]] 
    legend
} 
plot <-sample_plot_by_variant_by_day1(breath_mts_total_wide2, "av_quant_total_log10", "Total", TRUE,"e")
legend <- g_legend(plot)

```
combine the plots
```{r}
a <- sample_plot_by_variant_by_day1(breath_mts_total_wide2, "fine_log_quant", "Fine",FALSE,"a")
b <- sample_plot_by_variant(breath_mts_total_wide2,  "fine_log_quant", "Fine",compare_for_fine_and_total,label_fine_and_total,"b")
c <- sample_plot_by_variant_by_day1(breath_mts_total_wide2, "coarse_log_quant", "Coarse", FALSE,"c")
d <- sample_plot_by_variant(breath_mts_total_wide2, "coarse_log_quant", "Coarse",compare_for_coarse,label_coarse,"d")
e <- sample_plot_by_variant_by_day1(breath_mts_total_wide2, "av_quant_total_log10", "Total", FALSE,"e")
f <- sample_plot_by_variant(breath_mts_total_wide2, "av_quant_total_log10", "Total",compare_for_fine_and_total,label_fine_and_total,"f")
figure_2 <- grid.grab(wrap.grobs = TRUE)
```

```{r}
total<- grid.arrange(a, b,
                     c, d,
                     e, f,
                     legend,
                      ncol=5, nrow=4,
                      layout_matrix = rbind(c(1,1,2),
                                            c(3,3,4),
                                            c(5,5,6),
                                            c(7,7,7)),
                      heights = c(3,3,3,1))

#ggsave(total,filename="../output/figure_2_eba_by_variants_by_time.png",device="png",width=11,height=7.77,units="in")
#ggsave(total,filename="../output/tiff/figure_2_eba_by_variants_by_time.tiff",device="tiff",width=11,height=7.77,units="in", dpi=300)
ggsave(total,filename="../output/si_fig1_eba_variants_by_time_lte_5days.png",device="png",width=11,height=7.77,units="in")
```

