---
title: "summary_for_nucleocapsid_positive"
author: "Jianyu"
date: "4/21/2022"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(table1)
library(rvest)
library(ggpubr)
library(grid)
library(gridExtra)
set.seed(42)
```

```{r}
spr <- read.csv(file = "../working_files/nonpaired_pcr_sx_spr.csv")
demo <- readRDS(file = "../working_files/nonpaired_demohist.RDS")

spr_nuc_pos <- spr %>% filter(!(anti_nuc_igg_titer_1 == -1))
demo_nuc_pos <- demo %>% filter(!(anti_nuc_igg_titer_1 == -1))
```

#Q1. Did this group of people shed virus in aerosols?
```{r}
breath_samples <- c("G-II Coarse Aerosol","G-II Coarse Aerosol mask","G-II Fine Aerosol","G-II Fine Aerosol mask")
spr_breath <- spr_nuc_pos %>% filter(sample_type %in% breath_samples)
range(spr_breath$av_quant)

#conclusion: they didn't shed virus in aerosols.
```

#Q2. Did they report prior infections?  - Need to check the records manually or wait for their response in emails. 

#Q3. How many days since their symptom onset when they did the visit?
```{r}
spr_breath %>% filter(sample_type == "G-II Fine Aerosol mask") %>% group_by(subject_id) %>% summarise(min(days_past_onset),
                                                                                                      max(days_past_onset))
```

#Q4. What vaccine type they received? And how many days since their last shot?
```{r}
spr_nuc_pos[c('shot2_when1', 'shot2_when')] <- str_split_fixed(spr_nuc_pos$covid_shot_2_range, ';', 2)
spr_nuc_pos[c('shot1_when1', 'shot1_when')] <- str_split_fixed(spr_nuc_pos$covid_shot_1_range, ';', 2)
demo_nuc_pos[c('shot2_when1', 'shot2_when')] <- str_split_fixed(demo_nuc_pos$covid_shot_2_range, ';', 2)
demo_nuc_pos[c('shot1_when1', 'shot1_when')] <- str_split_fixed(demo_nuc_pos$covid_shot_1_range, ';', 2)

spr_nuc_pos <- spr_nuc_pos %>% 
  mutate(date_last_vax = if_else(booster == "boosted", booster_when, if_else(covid_shot_type %in% c("1", "2"), shot2_when1, shot1_when1))) %>% 
  mutate(days_since_vax = as.Date(sample_date) - as.Date(date_last_vax))
range(spr_nuc_pos$days_since_vax)

demo_nuc_pos <- demo_nuc_pos %>% 
  mutate(date_last_vax = if_else(booster == "boosted", booster_when, if_else(covid_shot_type %in% c("1", "2"), shot2_when1, shot1_when1)))

#summary of the booster status (not boosted vs. boosted), vaccine type, booster type, and when they received each of the shots
#type 1: Pfizer, 2. Moderna, C. J&J
demo_nuc_pos %>% select(subject_id, date_of_enrollment, booster, covid_shot_type, booster_type, booster_when, shot1_when1, shot2_when1, date_last_vax)

#time since their last shot when they did the visits
spr_nuc_pos %>%  group_by(subject_id) %>% summarise(min(days_since_vax), max(days_since_vax))
```

#Q5. What variants they had?
```{r}
table(demo_nuc_pos$seq_variant)
```

#Q6. Make a summary table for these cases. 
a summary table of the 5 anti-N IgG positive participants and their PCR and culture results for all sample types
Also in the table should be a column for booster and previous infections reported. 
```{r}
spr_nuc_pos$prior_infection[spr_nuc_pos$subject_id == "570"] <- "twice"
spr_nuc_pos$prior_infection[spr_nuc_pos$subject_id == "536"] <- "once"
spr_nuc_pos$prior_infection[spr_nuc_pos$subject_id == "583"] <- "never before"
spr_nuc_pos$prior_infection[spr_nuc_pos$subject_id == "571"] <- "never before"
spr_nuc_pos$prior_infection[spr_nuc_pos$subject_id == "551"] <- "not report"
summary <- spr_nuc_pos %>% select(subject_id, sample_date, seq_variant, sample_type, sample_quant, culture_status, booster, prior_infection, days_past_onset)
summary <- summary[order(summary$subject_id, summary$sample_date, summary$sample_type),]
samples <- c("G-II Coarse Aerosol", "G-II Fine Aerosol", "Midturbinate Swab", "Phone Swab", "Saliva")
summary2 <- summary %>% filter(sample_type %in% samples) %>% filter(!(culture_status=="not cultured"))

summary <- setNames(summary, c("Subject ID", "Sample Date", "Variant", "Sample Type", "qPCR", "Culture Status", "Booster Status", "Prior Infection", "Days post symptom onset"))
#write.csv(summary,file="../output/summary_5_cases.csv",row.names=F)


summary2 <- setNames(summary2, c("Subject ID", "Sample Date", "Variant", "Sample Type", "qPCR", "Culture Status", "Booster Status", "Prior Infection", "Days post symptom onset"))
#write.csv(summary2,file="../output/summary_5_cases_2.csv",row.names=F)
```

#Q7: Did these 5 participants have relatively lower viral loads in MTS than the rest of the study participants? 
```{r}
spr_omicon <- spr %>% filter(!(seq_variant == "Delta"))

stat_box_data <- function(y, upper_limit = 9.085) {
  data.frame(
    y = -0.02 * upper_limit,
    label = paste0('n=', length(y), '\n')
  )
}

anti_N_plot <- function(sample_name, yaxis, letter) {
  spr_sample <- spr_omicon %>% filter(sample_type == sample_name) %>% 
    mutate(nuc = as.factor(ifelse(anti_nuc_igg_titer_1=="-1", "Negative", "Positive")),
           log_quant = ifelse(sample_quant ==0, 0, log10(sample_quant)))
  
  anti_N_compare <- spr_sample %>% 
    ggplot(aes(x=nuc, y = log_quant))+
    geom_boxplot(aes(color = nuc),show.legend = FALSE)+
    geom_point(aes(color = nuc),show.legend = FALSE)+
    labs(x= "Anti-nucleocapsid IgG", 
         y= yaxis)+
    stat_compare_means(size = 4)+
    stat_summary(
      fun.data = stat_box_data, 
      geom = "text", 
      hjust = 0.5,
      vjust = 0.9,
      size=6,
      position = position_dodge(0.9)
    ) +
    theme_bw()+
    ggtitle(letter)+
    theme(axis.title=element_text(size=15, face='bold'),
          axis.text = element_text(size = 15),
          plot.title=element_text(size=20, face="bold"))
 return(anti_N_compare)
}


```

```{r}
blank <- grid.rect(gp=gpar(col="white"))
figure_e_5 <- grid.arrange(
  anti_N_plot("Midturbinate Swab", "MTS(log10)", "a"), 
  anti_N_plot("Saliva", "Saliva(log10)", "b"), 
  anti_N_plot("Phone Swab", "Fomite(log10)", "c"), 
  blank,
  anti_N_plot("G-II Fine Aerosol", "Fine EBA(log10)","d"), 
  anti_N_plot("G-II Coarse Aerosol", "Coarse EBA(log10)","e"), 
  layout_matrix = rbind(c(1,1,2,2,3,3),
                        c(4,4,4,4,4,4),
                        c(5,5,5,6,6,6)),
  heights = c(5,1,5)
)

figure_e_5
ggsave(figure_e_5,filename="../output/si_fig6_N-antigen_omicron.png",device="png",width=11,height=7.77,units="in")
```

