---
title: "table_1_total_93_cases"
author: "Jianyu"
date: "9/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(table1)
library(rvest)
library(car)
```

This script is used for creating table 1 for all 78 cases after we excluding those samples collected more than 5 days post symptom onset. 
Need to unmute or mute the filter lines for different groups.
```{r}
rpr_after <- read.csv(file = "../working_files/nonpaired_pcr_sx_rpr.csv")
rpr_after <- rpr_after %>% filter(days_past_onset<6|is.na(days_past_onset))
spr <- read.csv(file = "../working_files/nonpaired_pcr_sx_spr.csv")
spr <- spr %>% filter(days_past_onset<6|is.na(days_past_onset))

list2<-unique(spr$subject_id)
demo_after <- readRDS(file = "../working_files/nonpaired_demohist.RDS")
demo_after <- demo_after %>% filter((subject_id %in% list2)) 

rpr_before <- read.csv(file = "../working_files/manuscript_1_all_61_nonpaired_pcr_sx_rpr.csv")
rpr_before <- rpr_before %>% filter(days_past_onset<6|is.na(days_past_onset))
spr_before <- read.csv(file = "../working_files/manuscript_1_all_61_nonpaired_pcr_sx_spr.csv")
spr_before <- spr_before %>% filter(days_past_onset<6|is.na(days_past_onset))
list<-unique(spr_before$study_id)
demo_before <-readRDS(file="../working_files/manuscript_1_all_61_nonpaired_demohist.RDS")
demo_before <- demo_before %>% filter((study_id %in% list)) 

sgtf_samples <- spr_before %>% filter(S.gene.dropout)
sgtf_sids <- sgtf_samples$study_id %>% unique() %>% sort()
demo_before$sgtf_flag <- demo_before$study_id %>% sapply(function(x) {(x %in% sgtf_sids) %>% as.numeric()})

#for vaccine/booster - EBA tabulation
prior_fine_lod <- rpr_before %>% filter(any_rep_lod1)%>% 
  filter (sample_type == "G-II Fine Aerosol") %>% mutate(subject_id = as.factor(study_id)) %>% select (subject_id) %>% unique() %>% mutate(fine_pos = "TRUE")

prior_coarse_lod <- rpr_before %>% filter(any_rep_lod1) %>% 
  filter (sample_type == "G-II Coarse Aerosol") %>% mutate(subject_id = as.factor(study_id)) %>% select (subject_id) %>% unique()%>% mutate(coarse_pos = "TRUE")

latter_fine_lod <- rpr_after %>% filter(any_rep_lod1)%>% 
  filter (sample_type == "G-II Fine Aerosol") %>% mutate(subject_id = as.factor(subject_id)) %>% select (subject_id) %>% unique() %>% mutate(fine_pos = "TRUE")

latter_coarse_lod <- rpr_after %>% filter(any_rep_lod1) %>% 
  filter (sample_type == "G-II Coarse Aerosol") %>% mutate(subject_id = as.factor(subject_id)) %>% select (subject_id) %>% unique()%>% mutate(coarse_pos = "TRUE")

table(demo_after$seq_variant)

#additional dataset cleanup
demo_before$age_cat[demo_before$age_entry<18] <- "less than 18"
demo_before$age_cat[which(demo_before$age_entry>=18 & demo_before$age_entry<45)] <- "18 to 45"
demo_before$age_cat[demo_before$age_entry>=45] <- "greater than 45"
demo_before <- demo_before %>% 
  mutate(subject_id = as.factor(study_id),
         seq_variant = if_else(sgtf_flag=="1","Alpha","Other")) %>%
  #filter(period=="Negative") %>%
  mutate(booster = "not boosted",
         first = if_else(subject_id %in% c("482", "483", "484"), "1", "0"),
         full = "not",
         bmi = as.numeric(bmi),
         anti_rbd = ifelse(is.na(anti_rbd_igg_titer_1), NA,
                                  ifelse(anti_rbd_igg_titer_1==-1, "No","Yes")),
         anti_nuc = NA) %>% 
  select(subject_id, age_entry, sex_birth,seq_variant, booster, first, full, age_cat, race, other_race, latino, bmi, cri, covid_booster_actual,real_onset, anti_rbd, anti_nuc, covid_shot_type) 

demo_before <- demo_before %>% 
  left_join(prior_fine_lod, by = "subject_id") %>% 
  left_join(prior_coarse_lod, by = "subject_id")

demo_before <- demo_before %>% 
  mutate(fine_pos = if_else(is.na(fine_pos), "FALSE", fine_pos),
         coarse_pos = if_else(is.na(coarse_pos), "FALSE", coarse_pos),
         period = "Before",
         covid_shot_type = if_else(is.na(covid_shot_type),"0",covid_shot_type),
         booster_type = "0", # set to zero if NA
         covid_booster_actual = "0",
         symptomatic = if_else(is.na(real_onset), "No", "Yes"),
         latino = factor(latino, 
                          levels = c("0", "1", "9"),
                          labels = c("Not Hispanic", "Hispanic", "Decline")))

demo_after <- demo_after %>% 
  mutate(subject_id = as.factor(subject_id)) %>% 
  mutate(first ="1", 
         full = "yes",
         bmi = as.numeric(bmi),
         anti_rbd = "Yes",
         anti_nuc = ifelse(anti_nuc_igg_titer_1 == -1, "No", "Yes") ) %>% 
  select(subject_id, age_entry, sex_birth,seq_variant, booster, first, full, age_cat, race, other_race, latino, bmi, cri,covid_shot_type,booster_type, covid_booster_actual,real_onset, anti_rbd, anti_nuc) 

demo_after <- demo_after %>% 
  left_join(latter_fine_lod, by = c("subject_id")) %>% 
  left_join(latter_coarse_lod, by = c("subject_id")) %>% 
  mutate(fine_pos = if_else(is.na(fine_pos), "FALSE", fine_pos),
         coarse_pos = if_else(is.na(coarse_pos), "FALSE", coarse_pos),
         period = "After",
         cri = (cri==1),
         symptomatic = if_else(is.na(real_onset), "No", "Yes"),
         latino = factor(latino, 
                          levels = c("0", "1", "9"),
                          labels = c("Not Hispanic", "Hispanic", "Decline")))

demo <- bind_rows(demo_before,demo_after)
demo <- demo %>% mutate(seq_variant=factor(seq_variant,
                                           levels = c("Other", "Alpha", "Delta", "Omicron BA.1", "Omicron BA.1.1","Omicron BA.2"),
                                           labels = c("Other", "Alpha", "Delta", "Omicron BA.1", "Omicron BA.1.1","Omicron BA.2")))


spr_before <- spr_before %>% 
  mutate(subject_id = as.factor(study_id),
         period = "Before") %>% 
  select(subject_id, sample_date,sample_type, days_past_onset, cough_count, oral_temp, ox_sat, ur, lr, sys, gi, period)

spr_after <- spr %>% 
  mutate(subject_id = as.factor(subject_id),
         period = "After") %>% 
  select(subject_id, sample_date,sample_type,days_past_onset, cough_count, oral_temp, ox_sat,ur, lr, sys, gi,period)

spr_total <- bind_rows(spr_after,spr_before)

```

Vaccine
```{r}
demo <- demo %>% mutate(covid_shot_type=factor(covid_shot_type,
                                           levels = c("1", "2", "3", "4","0"),
                                           labels = c("BNT162b2", "mRNA-1273", "Ad26.COV2.S", "NVX-CoV2373","Not vaccinated")),
                        booster_type=factor(booster_type,
                                           levels = c("1", "2", "4", "0"),
                                           labels = c("BNT162b2", "mRNA-1273", "NVX-CoV2373", "Not boosted")))
demo$booster_type[is.na(demo$booster_type) | demo$covid_booster_actual == "0"] <- "Not boosted"

vax_booster_table <- table(demo$covid_shot_type, demo$booster_type)
```
Vaccination status
```{r}
demo$vacc_status <- demo %>% apply(1,function(x) {
  if (x["period"]=="After" & x["booster"]=="boosted") {
    y <- "Boosted"
  }
  if (x["period"]=="After" & x["booster"]=="not boosted") {
    y <- "Fully vaccinated, not boosted"
  }
  if (x["period"]=="Before" & x["covid_shot_type"]=="Not vaccinated") {
    y <- " Unvaccinated"
  }
  if (x["period"]=="Before" & x["covid_shot_type"]!="Not vaccinated") {
    y <- "Partially vaccinated"
  }
  return(y)
})
```


```{r}
table_vacc <- table1(~ fine_pos+
                        coarse_pos|full,
                      data=demo, overall="All subjects")

table_vacc
#write.csv(table_vacc,file="../output/table_vacc.csv",row.names=F)
table_booster <- table1(~ fine_pos+
                        coarse_pos|booster,
                      data=demo, overall="All subjects")

table_booster
#write.csv(table_booster,file="../output/table_booster.csv",row.names=F)
```


```{r}
label(demo$sex_birth) <- "Sex"
label(demo$age_entry) <-"Age"
label(demo$age_cat) <- "Age group"
label(demo$race) <- "Race"
label(demo$other_race) <- "Other race"
label(demo$bmi) <- "BMI"
label(demo$cri) <- "Chronic respiratory illness"
label(demo$seq_variant) <- "Variant"
label(demo$booster) <- "Booster"
```

table1
```{r}
table1_demo <- table1(~ seq_variant+
                        (sex_birth) +
                        age_entry +
                        (age_cat) +
                        race +
                        latino+
                        bmi +
                        as.character(cri) +
                        vacc_status+
                        covid_shot_type +
                        booster+
                        anti_rbd +
                        anti_nuc+
                        symptomatic
                      |period, 
                      data=demo, overall="All subjects")

table1_demo
```

```{r}
# since table1df has multiple rows that start with the same thing, use this cutoff function
cutofftable <- function(t,v) {
                startrow <- which(t$vars==v)
                toutput <- t[c(startrow:nrow(t)),]
                return(toutput)
}

# deal with sig figs: take "18.1 (36.7)" and return "18 (37)" if digits=0
parenformat <- function(x,d) {
                y1 <- x %>% strsplit(" \\(") %>% sapply(head,1) %>% as.numeric() %>% round(digits=d)
                 y2 <- x %>% strsplit(" \\(") %>% sapply(tail,1) %>% gsub("\\)","",.) %>% as.numeric() %>% round(digits=d)
                y <- paste0(y1," (",y2,")")
                return(y)
}

# similar function: take "25.93 ± 8.51" and return "25.9 ± 8.5"
plusminusformat <- function(x,d) {
                y1 <- x %>% strsplit(" ± ") %>% sapply(head,1) %>% as.numeric() %>% round(digits=d)
                y2 <- x %>% strsplit(" ± ") %>% sapply(tail,1) %>% as.numeric() %>% round(digits=d)
                y <- paste0(y1," ± ",y2)
                return(y)
}
```

```{r export the table 1 output}

table1df <- as.data.frame(read_html(table1_demo) %>% html_table(fill=T))
firstrow <- colnames(table1df) %>% gsub("[^0-9]","",.)

#names(firstrow) <- colnames(table1a)
table1a <- matrix(firstrow,ncol=4) %>% as.data.frame(stringsAsFactors=F)
colnames(table1a) <- c("vars","Enrolled September, 2021 - March, 2022","Enrolled June, 2020 - April 2021","All participants")
table1a$vars[1] <- "Number of participants"

colnames(table1df) <-  colnames(table1a)

# exhaled breath samples count comes from nonpaired pcr_dat file
negbreathcountdf <- spr_total %>% filter(period=="Before") %>% distinct(subject_id, sample_date, sample_type) %>% group_by(sample_type) %>% summarize(Cases = length(unique(subject_id)),`qPCR Samples` = n())
negbreathcount <- negbreathcountdf %>% filter(grepl("G-II",sample_type)) %>% pull(`qPCR Samples`) %>% max()

posbreathcountdf <- spr_total %>% filter(period=="After") %>% distinct(subject_id, sample_date, sample_type) %>% group_by(sample_type) %>% summarize(Cases = length(unique(subject_id)),`qPCR Samples` = n())
posbreathcount <- posbreathcountdf %>% filter(grepl("G-II",sample_type)) %>% pull(`qPCR Samples`) %>% max()
totalbreathcount <- negbreathcount+posbreathcount

ebsrow <- c("Number of exhaled breath samples",posbreathcount,negbreathcount,totalbreathcount)
names(ebsrow) <- colnames(table1a)
table1a <- table1a %>% bind_rows(ebsrow)

variantheader <- c("Variant, N(%)", "","","")
names(variantheader) <- colnames(table1a)
table1a <- table1a %>% bind_rows(variantheader)

variantlabels <- c("Other", "Alpha", "Delta","Omicron BA.1", "Omicron BA.1.1", "Omicron BA.2")
variant <- matrix(variantlabels,ncol=1) %>% as.data.frame(stringsAsFactors=F)
colnames(variant) <- "vars"
variant <- variant %>% left_join(table1df,by="vars")
variant <- variant %>% apply(c(1,2),function(x) {x %>% gsub("%","",.) %>% parenformat(0)}) %>% as.data.frame(stringsAsFactors=F)
variant$vars <- c("Other", "Alpha", "Delta","Omicron BA.1","Omicron BA.1.1","Omicron BA.2")
table1a <- table1a %>% bind_rows(variant)


femalerow <- table1df %>% filter(vars=="0") %>% head(1)
femalerow <- femalerow %>% gsub("%","",.) %>% parenformat(0)
names(femalerow) <-  colnames(table1a)
femalerow["vars"] <- "Female, N (%)"
table1a <- table1a %>% bind_rows(femalerow)

agemeanrow <- table1df %>% filter(vars=="Mean (SD)") %>% head(1)
agemeanrow <- agemeanrow %>% gsub(" \\("," ± ",.) %>% gsub("\\)","",.) %>% plusminusformat(1)
names(agemeanrow) <-  colnames(table1a)
agemeanrow["vars"] <- "Age, mean ± SD"
table1a <- table1a %>% bind_rows(agemeanrow)

agegroupheader <- c("Age group, N(%)", "", "", "")
names(agegroupheader) <- colnames(table1a)
table1a <- table1a %>% bind_rows(agegroupheader)

agegrouplabels <- c("less than 18","18 to 45","greater than 45")
agegroups <- matrix(agegrouplabels,ncol=1) %>% as.data.frame(stringsAsFactors=F)
colnames(agegroups) <- "vars"
agegroups <- agegroups %>% left_join(table1df,by="vars")
agegroups <- agegroups %>% apply(c(1,2),function(x) {x %>% gsub("%","",.) %>% parenformat(0)}) %>% as.data.frame(stringsAsFactors=F)
agegroups$vars <- c("<18","18-45",">45")
table1a <- table1a %>% bind_rows(agegroups)

raceheader <- c("Race/Ethnicity, N(%)", "", "", "")
names(raceheader) <- colnames(table1a)
table1a <- table1a %>% bind_rows(raceheader)

whiterow <- table1df %>% filter(grepl("White",vars)) %>% apply(c(1,2),function(x) {x %>% gsub("%","",.) %>% parenformat(0)})
names(whiterow) <- colnames(table1a)
whiterow["vars"] <- "White"
table1a <- table1a %>% bind_rows(whiterow)

blackrow <- table1df %>% filter(grepl("Black or African American",vars)) %>% apply(c(1,2),function(x) {x %>% gsub("%","",.) %>% parenformat(0)})
names(blackrow) <- colnames(table1a)
blackrow["vars"] <- "Black/African American"
table1a <- table1a %>% bind_rows(blackrow)


hiscutoff <- cutofftable(table1df,"latino")
hisrow <- table1df %>% filter(vars=="Hispanic") %>% head(1)%>% apply(c(1,2),function(x) {x %>% gsub("%","",.) %>% parenformat(0)})
names(hisrow) <-  colnames(table1a)
hisrow["vars"] <- "Hispanic"
table1a <- table1a %>% bind_rows(hisrow)

bmicutoff <- cutofftable(table1df,"BMI")
bmirow <- bmicutoff %>% filter(vars=="Mean (SD)") %>% head(1) %>% gsub(" \\("," ± ",.) %>% gsub("\\)","",.) %>% plusminusformat(1)
names(bmirow) <-  colnames(table1a)
bmirow["vars"] <- "BMI, mean ± SD"
table1a <- table1a %>% bind_rows(bmirow)

#cri row
cricutoff <- cutofftable(table1df,"as.character(cri)")
crirow <- cricutoff %>% filter(vars=="TRUE") %>% head(1) %>% apply(c(1,2),function(x) {x %>% gsub("%","",.) %>% parenformat(0)})
names(crirow) <-  colnames(table1a)
crirow["vars"] <- "Chronic respiratory illness, N (%)"
table1a <- table1a %>% bind_rows(crirow)

boosterheader <- c("Vaccination status, N(%)", "","","")
names(boosterheader) <- colnames(table1a)
table1a <- table1a %>% bind_rows(boosterheader)
boosterlabels <- c("Boosted","Fully vaccinated, not boosted","Partially vaccinated", "Unvaccinated")
booster <- matrix(boosterlabels,ncol=1) %>% as.data.frame(stringsAsFactors=F)
colnames(booster) <- "vars"
booster <- booster %>% left_join(table1df,by="vars")
booster <- booster %>% apply(c(1,2),function(x) {x %>% gsub("%","",.) %>% parenformat(0)}) %>% as.data.frame(stringsAsFactors=F)
booster$vars <- c("Boosted","Fully vaccinated, not boosted","Partially vaccinated", "Unvaccinated")
table1a <- table1a %>% bind_rows(booster)

vaxheader <- c("Vaccine type, N(%)", "", "", "")
names(vaxheader) <- colnames(table1a)
table1a <- table1a %>% bind_rows(vaxheader)
vaxlabels <- c("BNT162b2","mRNA-1273", "Ad26.COV2.S", "NVX-CoV2373","Not vaccinated")
vax <- matrix(vaxlabels,ncol=1) %>% as.data.frame(stringsAsFactors=F)
colnames(vax) <- "vars"
vax <- vax %>% left_join(table1df,by="vars")
vax <- vax %>% apply(c(1,2),function(x) {x %>% gsub("%","",.) %>% parenformat(0)}) %>% as.data.frame(stringsAsFactors=F)
vax$vars <- c("BNT162b2","mRNA-1273", "Ad26.COV2.S", "NVX-CoV2373","Not vaccinated")
table1a <- table1a %>% bind_rows(vax)

#rbd
rbdcutoff <- cutofftable(table1df,"anti_rbd")
rbdrow <- rbdcutoff %>% filter(vars=="Yes") %>% head(1) %>% apply(c(1,2),function(x) {x %>% gsub("%","",.) %>% parenformat(0)})
names(rbdrow) <-  colnames(table1a)
rbdrow["vars"] <- "Anti-spike RBD antibody (IgG), N (%)"
table1a <- table1a %>% bind_rows(rbdrow)

#placeholder
nuccutoff <- cutofftable(table1df,"anti_nuc")
nucrow <- nuccutoff %>% filter(vars=="Yes") %>% head(1) %>% apply(c(1,2),function(x) {x %>% gsub("%","",.) %>% parenformat(0)})
names(nucrow) <-  colnames(table1a)
nucrow["vars"] <- "Anti-nucleocapsid antibody (IgG), N (%)"
table1a <- table1a %>% bind_rows(nucrow)


symcutoff <- cutofftable(table1df,"symptomatic")
symrow <- symcutoff %>% filter(vars=="Yes") %>% head(1) %>% apply(c(1,2),function(x) {x %>% gsub("%","",.) %>% parenformat(0)})
names(symrow) <-  colnames(table1a)
symrow["vars"] <- "Ever symptomatic, N (%)"
table1a <- table1a %>% bind_rows(symrow)
```

Separate section to summarize symptoms
This aggregates all reported scores on all the days exhaled breath samples were obtained
```{r}
breath_samples <- c("G-II Coarse Aerosol","G-II Coarse Aerosol mask","G-II Fine Aerosol","G-II Fine Aerosol mask")

#filter breath samples
sx_dat <- spr_total %>% filter(sample_type %in% c(breath_samples)) %>% filter(!is.na(days_past_onset))
sx_dat_period <- list()
sx_dat_period[["After"]] <- sx_dat %>% filter(period=="After")
sx_dat_period[["Before"]] <- sx_dat %>% filter(period=="Before")
sx_dat_period[["all"]] <- sx_dat %>% filter(period %in% c("After","Before"))

sx_dat_period <- sx_dat_period %>% lapply(function(x) {
                x %>% distinct(subject_id,sample_date,.keep_all=T)
})


summaryvars <- c("days_past_onset","cough_count","oral_temp","ox_sat","ur","lr","sys","gi")

sx_period_summary <- summaryvars %>% lapply(function(y) {
                sx_dat_period %>% sapply(function(x) {
                                x %>% summarise(
                                                av1=mean(eval(parse(text=y)),na.rm=T),
                                                sd1=sd(eval(parse(text=y)),na.rm=T),
                                                min1=min(eval(parse(text=y))),
                                                max1=max(eval(parse(text=y))),
                                                median1=median(eval(parse(text=y))),
                                                q1=quantile(eval(parse(text=y)),0.25,na.rm=T),
                                                q2=quantile(eval(parse(text=y)),0.75,na.rm=T))
                                })
})
names(sx_period_summary) <- summaryvars
sx_period_summary <- sx_period_summary %>% lapply(as.data.frame,stringsAsFactors=F)

tastesmelltable <- sx_dat_period %>% lapply(function(x) {table(x$taste_smell_loss)})
tastesmellpercentage <- tastesmelltable %>% lapply(function(x) {
                x["1"]/(sum(x))
}) %>% unlist() %>% `*`(100) %>% round(0) %>% paste0("%")
```
Make the table into a manuscript ready format
```{r}
table1b <- data.frame()
# start with the day past symptom onset row for table 1b
dporowmean <- sx_period_summary[["days_past_onset"]]["av1",] %>% unlist() %>% round(0)
dporowsd <- sx_period_summary[["days_past_onset"]]["sd1",]  %>% unlist() %>% round(0)
dporowmin <- sx_period_summary[["days_past_onset"]]["min",] %>% unlist %>% round(0)
dporowmax <- sx_period_summary[["days_past_onset"]]["max",] %>% unlist %>% round(0)
dporow <- mapply(FUN=function(x,y,z,w) {paste0(x," ± ",y," (",z,"-",w,")")},
                   x=dporowmean, y=dporowsd, z=dporowmin, w=dporowmax) %>% c("days_past_onset",.)

names(dporow) <- c("vars",names(sx_dat_period))
table1b <- table1b %>% bind_rows(dporow)


coughrowmean <- sx_period_summary[["cough_count"]]["av1",] %>% unlist() %>% round(0)
coughrowsd <- sx_period_summary[["cough_count"]]["sd1",]  %>% unlist() %>% round(0)
coughrowmin <- sx_period_summary[["cough_count"]]["min",] %>% unlist %>% round(0)
coughrowmax <- sx_period_summary[["cough_count"]]["max",] %>% unlist %>% round(0)
coughrow <- mapply(FUN=function(x,y,z,w) {paste0(x," ± ",y," (",z,"-",w,")")},
                   x=coughrowmean, y=coughrowsd, z=coughrowmin, w=coughrowmax) %>% c("cough_count",.)

names(coughrow) <- c("vars",names(sx_dat_period))
table1b <- table1b %>% bind_rows(coughrow)

symptable <- data.frame()
symptablelist <- c("ur","lr","sys","gi")
for (i in symptablelist) {
  imedian <- sx_period_summary[[i]]["median1",] %>% unlist %>% round(1)
  iq1 <- sx_period_summary[[i]]["q1",] %>% unlist %>% round(1)
  iq2 <- sx_period_summary[[i]]["q2",] %>% unlist %>% round(1)
  irow <- mapply(FUN = function(x,y,z) {paste0(x," (",y," - ",z,")")},
                 x=imedian, y=iq1, z=iq2) %>% c(i,.)
  names(irow) <- colnames(table1b)
  symptable <- symptable %>% bind_rows(irow)
}
table1b <- table1b %>% bind_rows(symptable)

temprowmean <- sx_period_summary[["oral_temp"]]["av1",] %>% unlist() %>% round(1)
temprowsd <- sx_period_summary[["oral_temp"]]["sd1",] %>% unlist() %>% round(1)
temprow <- mapply(FUN=function(x,y) {paste0(x," ± ",y)},
                   x=temprowmean, y=temprowsd) %>% c("temp",.)
names(temprow) <- colnames(table1b)
table1b <- table1b %>% bind_rows(temprow)

oxsatrowmean <- sx_period_summary[["ox_sat"]]["av1",] %>% unlist() %>% round(1)
oxsatrowsd <- sx_period_summary[["ox_sat"]]["sd1",] %>% unlist() %>% round(1)
oxsatrow <- mapply(FUN=function(x,y) {paste0(x," ± ",y)},
                   x=oxsatrowmean, y=oxsatrowsd) %>% c("oxsat",.)
names(oxsatrow) <- colnames(table1b)
table1b <- table1b %>% bind_rows(oxsatrow)

colnames(table1b) <- c("vars","Enrolled September, 2021 - March, 2022","Enrolled June, 2020 - April 2021","All participants")
table1b$vars <- c("Days post symptom onset, mean± SD (range)","Coughs per 30 min, mean± SD (range)","Median upper respiratory symptoms (IQR)","Median lower respiratory symptoms (IQR)","Median systemic symptoms (IQR)","Median gastrointestinal symptoms (IQR)","Temperature (C), mean ± SD","Oxygen saturation (SpO2), mean ± SD")
```
bind and write tables
```{r}
table_1_formatted <- bind_rows(table1a,table1b)
table_1_formatted <- table_1_formatted[, c("vars","Enrolled June, 2020 - April 2021","Enrolled September, 2021 - March, 2022","All participants")]

write.csv(table_1_formatted,file="../output/si_tab2_demo_lte_5days.csv",row.names=F)
```

Table 1 comparison
comparison of participant characteristics - demographics
```{r eval=FALSE, include=FALSE}
demo <- demo %>% mutate(period = as.factor(period))
table1c <- matrix(ncol=3, nrow=0) %>% as.data.frame(stringsAsFactors=F)
###T-test
ttest_list <- c("age_entry", "bmi")
for (i in ttest_list){
  lttest <- leveneTest(eval(parse(text=i)) ~ period, data = demo)
  if (lttest$`Pr(>F)`[1] >= 0.05) {
    ttest_vars <- t.test(eval(parse(text=i)) ~ period, data = demo, var.equal= T)
  }
  else{
    ttest_vars<-t.test(eval(parse(text=i)) ~ period, data = demo, var.equal= F)
  }
  if (ttest_vars$p.value >=0.05){
    note <- "-"
  }
  else{
    note <- "*"
  }
  table1c <- table1c %>% rbind(c(i,ttest_vars$p.value,note))
}

###Chi-square test
chisq_list <- c("sex_birth", "race", "latino","cri",)
for (i in chisq_list){
  chitest <- chisq.test(demo[,i], demo$period)
  if(chitest$p.value >=0.05){
    note <- "-"
  }
  else{
    note <- "*"
  }
  table1c <- table1c %>% rbind(c(i,chitest$p.value,note))
}

###Fisher's exact
fish_list <- c("seq_variant","age_cat",  "vacc_status","anti_rbd","anti_nuc","symptomatic")
for (i in fish_list){
  fishtest <- fisher.test(demo[,i], demo$period)
  if(fishtest$p.value >=0.05){
    note <- "-"
  }
  else{
    note <- "*"
  }
  table1c <- table1c %>% rbind(c(i,fishtest$p.value,note))
}
colnames(table1c) <- c("Var", "p-value", "less than 0.05? (*)")
table1c 
#write.csv(table1c,file="../output/table_1_comparison_values.csv",row.names=F)
```

comparison of participant characteristics - symptoms
(The below lines don't seem to be right. Will need to update the coding to create the symptom comparison, especially for the oral temp)
```{r eval=FALSE, include=FALSE}
data_compare <- as.data.frame(sx_dat_period[["all"]]) %>% mutate(period = factor(period))
data_before <- as.data.frame(sx_dat_period[["Before"]])
data_after <- as.data.frame(sx_dat_period[["After"]])

vars2 <- c("days_past_onset","cough_count","ur","lr","sys","gi","oral_temp","ox_sat")
aggregate_allvars <- vars2 %>% lapply(function(x) {
  aggregate(eval(parse(text=x)) ~ period, data = data_compare, sd)
})
names(aggregate_allvars) <- vars2

for (i in 1:length(aggregate_allvars)) {
  colnames(aggregate_allvars[[i]])[colnames(aggregate_allvars[[i]])=="eval(parse(text = x))"] <- vars2[i]
}
aggregate_df <- aggregate_allvars[[1]]

for (i in 2:length(aggregate_allvars)) {
  aggregate_df <- aggregate_df %>% left_join(aggregate_allvars[[i]],by="period")
}
aggregate_df$test <- aggregate_df$period %>% paste0("aggregate.",.)

homo_vars <- vars2 %>% lapply(function(x) {
  leveneTest(eval(parse(text=x)) ~ period, data = data_compare)
})
names(homo_vars) <- vars2
homo_vars_pr_f <- homo_vars %>% sapply(function(x) {x %>% select(`Pr(>F)`) %>% filter(!is.na(`Pr(>F)`))}) %>% bind_rows()
colnames(homo_vars_pr_f) <- colnames(homo_vars_pr_f) %>% gsub("\\.Pr\\(>F\\)","",.)
homo_vars_pr_f$test <- "homo.Pr(>F)"

ttest_vars <- vars2 %>% sapply(function(x) {
  t.test(eval(parse(text=x)) ~ period, data = data_compare,var.equal= T) %>% `[[`("p.value")
})
ttest_vars$test <- "ttest.pvalue"
alltests2 <- bind_rows(aggregate_df,homo_vars_pr_f,ttest_vars) %>% select(test,all_of(vars2))

alltests2[,vars2] <- alltests2[,vars2] %>% apply(c(1,2),signif,2)
```