---
title: "model_omicron"
author: "Jianyu"
date: "3/5/2022"
output:
  html_document: default
---
analysis of shedding/variants
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Matrix)
library(mvtnorm)
library(lme4)
library(lmerTest)
library(tidyverse)
source("z_create_matrix.R")
set.seed(42)
```

Data: nomask rpr and spr files, filtered for seronegative participants
```{r}
rpr <- read.csv("../working_files/nomask_pcr_sx_rpr.csv")
spr <- read.csv("../working_files/nomask_pcr_sx_spr.csv")

#include only omicron case
rpr <- rpr %>% filter(!(seq_variant=="Delta"))
spr <- spr %>% filter(!(seq_variant=="Delta"))
```


create a variable "days_since_vax" for Days since last vaccine or booster
```{r}
spr[c('shot2_when1', 'shot2_when')] <- str_split_fixed(spr$covid_shot_2_range, ';', 2)
rpr[c('shot2_when1', 'shot2_when')] <- str_split_fixed(rpr$covid_shot_2_range, ';', 2)
spr[c('shot1_when1', 'shot1_when')] <- str_split_fixed(spr$covid_shot_1_range, ';', 2)
rpr[c('shot1_when1', 'shot1_when')] <- str_split_fixed(rpr$covid_shot_1_range, ';', 2)


spr <- spr %>% 
  mutate(anti_nuc = if_else(anti_nuc_igg_titer_1 == -1, 0, 1),
         boosted = if_else(booster == "boosted", 1, 0)) %>% 
  mutate(date_last_vax = if_else(booster == "boosted", booster_when, if_else(covid_shot_type %in% c("1", "2"), shot2_when1, shot1_when))) %>% 
  mutate(days_since_vax = as.Date(sample_date) - as.Date(date_last_vax),
         BA2 = if_else(seq_variant=="Omicron BA.2", 1, 0))

rpr <- rpr %>% 
  mutate(anti_nuc = if_else(anti_nuc_igg_titer_1 == -1, 0, 1),
         boosted = if_else(booster == "boosted", 1, 0)) %>%
  mutate(date_last_vax = if_else(booster == "boosted", booster_when, if_else(covid_shot_type %in% c("1", "2"), shot2_when1, shot1_when))) %>%  
  mutate(days_since_vax = as.Date(sample_date) - as.Date(date_last_vax),
         BA2 = if_else(seq_variant=="Omicron BA.2", 1, 0))

range(spr$days_since_vax) #8-430

nrow(spr %>% filter(anti_nuc == 1) %>% select(subject_id) %>% unique()) #5
```

spr to subset the sample_quant values (at sample level) for mt and saliva that will be used as predictors for later analysis
```{r}
mt <- spr %>% filter(sample_type == "Midturbinate Swab")
mt <- mt %>% mutate(mt_av_quant = av_quant,
                    log_mt = if_else(mt_av_quant==0, 0, log(mt_av_quant))) 
mt <- mt %>% select(log_mt, mt_av_quant, subject_id, sample_date)

saliva <- spr %>% filter(sample_type == "Saliva") 
saliva <- saliva %>% mutate(sal_av_quant = av_quant,
                             log_sal = if_else(sal_av_quant==0, 0, log(sal_av_quant)))
saliva <- saliva %>% select(log_sal, sal_av_quant, subject_id, sample_date)
```
Use left_join on rpr to add mt and saliva values (at sample level) to the three datasets so we can use these variables as predictors for later analysis
```{r}
fine <- rpr %>% filter(sample_type == "G-II Fine Aerosol") %>%
  left_join(mt, by=c("subject_id"="subject_id", "sample_date"="sample_date")) %>%
  left_join(saliva, by=c("subject_id"="subject_id", "sample_date"="sample_date")) %>% 
  mutate(yL_fine = log(sample_quant+1))
length(unique(fine$subject_id))
length(unique(fine$sample_id))

coarse <- rpr %>% filter(sample_type == "G-II Coarse Aerosol") %>%
  left_join(mt, by=c("subject_id"="subject_id", "sample_date"="sample_date")) %>% 
  left_join(saliva, by=c("subject_id"="subject_id", "sample_date"="sample_date")) %>% 
  mutate(yL_coarse = log(sample_quant+1))

length(unique(coarse$subject_id))
length(unique(coarse$sample_id))
```

##fine aerosol lmer

BIVARIATE ANALYSIS FOR FINE AEROSOL
```{r}
est_ci_fine_list <- list()

vars_fine <- c( 'BA2', 'boosted','anti_nuc','days_since_vax', 'age_entry','sex_birth','days_past_onset', 'log_mt', 'log_sal', 'cough_count', 'ur', 'lr', 'gi', 'sys','boosted*BA2','boosted*age_entry','boosted*sex_birth')

for (i in vars_fine) {
  print(paste0("BIVARIATE ANALYSIS FOR FINE AEROSOL AND ",i))
  lm <- lmer (yL_fine ~ (1|subject_id) + (1|subject_id:sample_date) + eval(parse(text=i)), data=fine)
  est <- coef(summary(lm))[2]
  se <- sqrt(diag(vcov(lm)))[2]
  up <- est+1.96*se
  low <- est-1.96*se
  est_ci_fine_list[[i]] <- c(i, est, low, up)
  names(est_ci_fine_list[[i]]) <- c("varname","beta_lmer","ci_low","ci_high")
}
est_ci_fine_list_df <- bind_rows(est_ci_fine_list)

```

check which variables go into the next selection: 90% CI: est +/- 1.645*se
```{r eval=FALSE, include=FALSE}
est_ci_fine_list2 <- list()

vars_fine <- c( 'BA2', 'boosted','anti_nuc','days_since_vax', 'age_entry','sex_birth','days_past_onset', 'log_mt', 'log_sal', 'cough_count', 'ur', 'lr', 'gi', 'sys','boosted*BA2','boosted*age_entry','boosted*sex_birth')


for (i in vars_fine) {
  print(paste0("BIVARIATE ANALYSIS FOR FINE AEROSOL AND ",i))
  lm <- lmer (yL_fine ~ (1|subject_id) + (1|subject_id:sample_date) + eval(parse(text=i)), data=fine)
  est <- coef(summary(lm))[2]
  se <- sqrt(diag(vcov(lm)))[2]
  up <- est+1.645*se
  low <- est-1.645*se
  est_ci_fine_list2[[i]] <- c(i, est, low, up)
  names(est_ci_fine_list2[[i]]) <- c("varname","beta_lmer","ci_low","ci_high")
}
est_ci_fine_list_df2 <- bind_rows(est_ci_fine_list2)

#anti_nuc, age, log_mt, log_sal, cough_count, sys
```

##model selection based on AIC: 
full model: ~age_entry+sex_birth+log_mt+log_sal+cough_count+sys
keep: ~age_entry+sex_birth
```{r eval=FALSE, include=FALSE}
vars <- c( "anti_nuc",'log_mt', 'log_sal', 'cough_count', 'sys')

models <- list()

for (i in 1:5){
  vc <- combn(vars,i)
  for (j in 1:ncol(vc)){
    model <- paste0("BA2+age_entry+sex_birth+", paste0(vc[,j], collapse = "+"))
    models <- c(models, model)
  }
}

aic_final = 100000

for (i in models){
  lm <- lmer (as.formula(paste0("yL_fine ~(1|subject_id) + (1|subject_id:sample_date) +", i)), data=fine)
aic <- summary(lm)$AIC

if (aic_final>aic){
  aic_final <- aic
  model_final <- i
}
print(i)
}
aic_final #248.2216 
model_final #BA2+age_entry+sex_birth+anti_nuc+log_sal+cough_count+sys
```

Multivariable analysis for Fine aerosol
```{r}
est_ci_fine_multi <- c()
lm <- lmer (yL_fine ~ (1|subject_id) + (1|subject_id:sample_date) + BA2+age_entry+sex_birth+anti_nuc+log_sal+cough_count+sys , data=fine)

vars <- c('intercept', 'BA2',"age_entry", "sex_birth","anti_nuc","log_sal", "cough_count", "sys")
for (i in c(2:length(vars))){
                est <- coef(summary(lm))[i]
                se <- sqrt(diag(vcov(lm)))[i]
                up <- est+1.96*se
                low <- est-1.96*se
                est_ci <- c(vars[i], est, low, up)
                est_ci_fine_multi <- rbind(est_ci_fine_multi, est_ci)
}
colnames(est_ci_fine_multi) <- c("varname","beta_lmer","ci_low","ci_high")

```

coarse aerosol lmer

BIVARIATE ANALYSIS FOR COARSE AEROSOL
```{r}
est_ci_coarse_list <- list()

vars_coarse <- c( 'BA2', 'boosted','anti_nuc','days_since_vax', 'age_entry','sex_birth','days_past_onset', 'log_mt', 'log_sal', 'cough_count', 'ur', 'lr', 'gi', 'sys','boosted*BA2','boosted*age_entry','boosted*sex_birth')


for (i in vars_coarse) {
  print(paste0("BIVARIATE ANALYSIS FOR COARSE AEROSOL AND ",i))
  lm <- lmer (yL_coarse ~ (1|subject_id) + (1|subject_id:sample_date) + eval(parse(text=i)), data=coarse)
  est <- coef(summary(lm))[2]
  se <- sqrt(diag(vcov(lm)))[2]
  up <- est+1.96*se
  low <- est-1.96*se
  est_ci_coarse_list[[i]] <- c(i, est, low, up)
  names(est_ci_coarse_list[[i]]) <- c("varname","beta_lmer","ci_low","ci_high")
}
est_ci_coarse_list_df <- bind_rows(est_ci_coarse_list)
```

check which variables go into the next selection: 90% CI: est +/- 1.645*se
```{r eval=FALSE, include=FALSE}
est_ci_coarse_list2 <- list()


vars_coarse <- c( 'BA2', 'boosted','anti_nuc','days_since_vax', 'age_entry','sex_birth','days_past_onset', 'log_mt', 'log_sal', 'cough_count', 'ur', 'lr', 'gi', 'sys','boosted*BA2','boosted*age_entry','boosted*sex_birth')


for (i in vars_coarse) {
  print(paste0("BIVARIATE ANALYSIS FOR COARSE AEROSOL AND ",i))
  lm <- lmer (yL_coarse ~ (1|subject_id) + (1|subject_id:sample_date) + eval(parse(text=i)), data=coarse)
  est <- coef(summary(lm))[2]
  se <- sqrt(diag(vcov(lm)))[2]
  up <- est+1.645*se
  low <- est-1.645*se
  est_ci_coarse_list2[[i]] <- c(i, est, low, up)
  names(est_ci_coarse_list2[[i]]) <- c("varname","beta_lmer","ci_low","ci_high")
}
est_ci_coarse_list_df2 <- bind_rows(est_ci_coarse_list2)

#boosted, anti_nuc, age_entry, days_past_onset, log_mt, log_sal, sys, boosted*age_entry, boosted*sex_birth
```
##model selection based on AIC: 
#keep: ~BA+age_entry+sex_birth
#could not include only the interaction terms without include the component variables, so the selection is among: 
#boosted+boosted*age_entry, boosted+boosted*sex_birth, boosted, days_past_onset, log_mt, log_sal, gi, sys
```{r eval=FALSE, include=FALSE}
vars <- c('boosted+boosted*age_entry','boosted+boosted*sex_birth','boosted','anti_nuc','days_past_onset', 'log_mt', 'log_sal', 'sys')

models <- list()

for (i in 1:8){
  vc <- combn(vars,i)
  for (j in 1:ncol(vc)){
    model <- paste0("BA2+age_entry+sex_birth+", paste0(vc[,j], collapse = "+"))
    models <- c(models, model)
  }
}

aic_final = 100000

for (i in models){
  lm <- lmer (as.formula(paste0("yL_coarse ~(1|subject_id) + (1|subject_id:sample_date) +", i)), data=coarse)
aic <- summary(lm)$AIC

if (aic_final>aic){
  aic_final <- aic
  model_final <- i
}
}
aic_final 
model_final 
#"BA2+age_entry+sex_birth+boosted+boosted*sex_birth+anti_nuc+log_sal+sys" 322.597
```

Multivariable analysis for coarse aerosol
```{r}
est_ci_coarse_multi <- c()
lm <- lmer (yL_coarse ~ (1|subject_id) + (1|subject_id:sample_date) + BA2+age_entry+sex_birth+boosted+anti_nuc+log_sal+sys+boosted*sex_birth, data=coarse)

vars <- c('intercept', 'BA2',"age_entry","sex_birth", 'boosted','anti_nuc',"log_sal","sys","boosted*sex_birth")
for (i in c(2:length(vars))){
                est <- coef(summary(lm))[i]
                se <- sqrt(diag(vcov(lm)))[i]
                up <- est+1.96*se
                low <- est-1.96*se
                est_ci <- c(vars[i], est, low, up)
                est_ci_coarse_multi <- rbind(est_ci_coarse_multi, est_ci)
}
colnames(est_ci_coarse_multi) <- c("varname","beta_lmer","ci_low","ci_high")
print(est_ci_coarse_multi)

aic <- summary(lm)$AIC
aic #322.597
```


create IQR adjustment values
these are multiplied into the power component of an exponential, so essentially 1 = no adjustment
these are needed for table 3, so save to working files
```{r}
# MT: IQR for the log value
mt_rpr <- rpr %>% filter(sample_type == "Midturbinate Swab") %>% 
  mutate(log_mt_rpr = if_else(sample_quant==0, 0, log(sample_quant)))
log_mt_IQR <- IQR(mt_rpr$log_mt_rpr) 

# saliva: IQR for the log value
sal_rpr <- rpr %>% filter(sample_type == "Saliva") %>% 
  mutate(log_sal_rpr = if_else(sample_quant==0, 0, log(sample_quant)))
log_sal_IQR <- IQR(sal_rpr$log_sal_rpr) 

#IQR for the symptom scores:  
breath_samples <- c("G-II Coarse Aerosol","G-II Coarse Aerosol mask","G-II Fine Aerosol","G-II Fine Aerosol mask")
#filter breath samples
sx_dat <- spr %>% filter(sample_type %in% c(breath_samples))

#IQRs for symptom scores were from table 1
IQRadjustlist <- list()
IQRadjustlist[["log_sal"]] <- log_sal_IQR
IQRadjustlist[["log_mt"]] <- log_mt_IQR
IQRadjustlist[["cough_count"]] <- 1
IQRadjustlist[["age_entry"]] <- 10
IQRadjustlist[["days_past_onset"]] <- 1
IQRadjustlist[["gi"]] <- IQR(sx_dat$gi)
IQRadjustlist[["lr"]] <- IQR(sx_dat$lr)
IQRadjustlist[["sys"]] <- IQR(sx_dat$sys)
IQRadjustlist[["ur"]] <- IQR(sx_dat$ur)
IQRadjustlist[["sex_birth"]] <- 1
IQRadjustlist[["boosted"]] <- 1
IQRadjustlist[["days_since_vax"]] <- 1
IQRadjustlist[["BA2"]] <- 1
IQRadjustlist[["boosted"]] <- 1
IQRadjustlist[["age_entry"]] <- 1
```

```{r}
exptransform <- function(x) {
  y2 <- exp(x) %>% signif(2)
  y3 <- exp(x) %>% signif(3)
  if(y2==1){y=y3}
  else(y=y2)
  return(y)
}

# fine/coarse unadjusted/unadjusted mean & CIs

meanCIlist <- list(est_ci_fine_list_df,est_ci_fine_multi,est_ci_coarse_list_df,est_ci_coarse_multi)
names(meanCIlist) <- c("fine_unadj","fine_adj","coarse_unadj","coarse_adj")

meanCIlist <- meanCIlist %>% lapply(as.data.frame,stringsAsFactors=F)
for (i in names(meanCIlist)) {
  meanCIlist[[i]]$iqradjust <- meanCIlist[[i]]$varname %>% lapply(function(x) {IQRadjustlist[[x]]}) %>% lapply(function(x) {if (length(x)==0) {1} else {x}}) %>% unlist()
  meanCIlist[[i]]$beta_1 <- ((meanCIlist[[i]]$beta_lmer %>% as.numeric())*meanCIlist[[i]]$iqradjust) %>% sapply(exptransform)
  meanCIlist[[i]]$ci_low_1 <- ((meanCIlist[[i]]$ci_low %>% as.numeric())*meanCIlist[[i]]$iqradjust) %>% sapply(exptransform)
  meanCIlist[[i]]$ci_high_1 <- ((meanCIlist[[i]]$ci_high %>% as.numeric())*meanCIlist[[i]]$iqradjust) %>% sapply(exptransform)
  meanCIlist[[i]]$predict1 <- meanCIlist[[i]][,c("beta_1","ci_low_1","ci_high_1")] %>% apply(1,function(x) {paste0(x["beta_1"]," (",x["ci_low_1"],", ",x["ci_high_1"],")")})
}


# for unadjusted models, don't need to adjusted entries
meanCIlist[["fine_unadj"]] <- meanCIlist[["fine_unadj"]] %>% filter(!grepl("adjusted",varname))
meanCIlist[["coarse_unadj"]] <- meanCIlist[["coarse_unadj"]] %>% filter(!grepl("adjusted",varname))

table3output <- meanCIlist[["coarse_unadj"]] %>% select(varname,predict1) %>% rename(`Coarse Unadjusted`=predict1)
table3output <- table3output %>% left_join(meanCIlist[["coarse_adj"]] %>% select(varname,predict1),by="varname") %>% rename(`Coarse Adjusted`=predict1)
table3output <- table3output %>% left_join(meanCIlist[["fine_unadj"]] %>% select(varname,predict1),by="varname") %>% rename(`Fine Unadjusted`=predict1)
table3output <- table3output %>% left_join(meanCIlist[["fine_adj"]] %>% select(varname,predict1),by="varname") %>% rename(`Fine Adjusted`=predict1)
table3output[is.na(table3output)] <- "-"

tableorder <- c('BA2','boosted','anti_nuc','days_since_vax', 'age_entry','sex_birth','days_past_onset', 'log_mt', 'log_sal', 'cough_count', 'ur', 'lr', 'gi', 'sys','boosted*BA2','boosted*age_entry','boosted*sex_birth')

table3output1 <- matrix(tableorder,ncol=1) %>% as.data.frame(stringsAsFactors=F)
colnames(table3output1) <- "varname"
table3output1 <- table3output1 %>% left_join(table3output,by="varname")

table3output1

write.csv(table3output1,file="../output/table_model_omicron_lmer.csv",row.names=F)
```
